DeepSphere: a graph-based spherical CNN

### 2019-09-18 main.tex
fixed third section

% The plan for this section
% 0) Explain why the Laplacian is so important...
By definition, the graph convolution (on the sphere) is rotation equivariant to rotation, if and only if, it commutes with the rotation operator. 
As, in the context of this work, graph convolution can be performed is performed by successive application of the graph Laplacian, i.e: $g(\b{L}) \b{f} = \left(\sum_{i=0}^K \alpha_i \b{L}^i\right) \b{f}$, if the graph Laplacian $\b{L}$ commutes with the rotation operator, then, by recursion, it will also commute with the convolution $g(\b{L})$.
Hence to simplify our analysis, we concentrate on the graph Laplacian.

% 0.5) Explain the sampling and metric
We first need to deal with the sampling problem. The rotation operator $R_{b}$ \nati{define rotation} act on the continuous signal $R(b), b\in SO(3): R(b)f(\omega) = f(g^{-1}\omega)$
graph laplacian on a discrete graph that is build from a sampling.
So we need a sampling operator, ideally invertible. \nati{Define sampling}
Eventually, we formulate the commute error of the graph Laplacian and the rotation operator as: 
equ...
Of course result will depend on the sampling and on the graph construction

% !) summary the work of Khasanova

% 2) explain our approach
Our approach is different as we do not miminize eq... Instead we inspire ourselves with Belkin, which says that for \emph{uniform sampling} the graph Laplacian converges toward the Laplace-Beltrami. b) The Laplace Beltrami commute with rotation.
The following theorem extend the work of Belkin and Niogi to fixed sampling \nati{(Including Healpix or not?)}
THEOREM HERE
This is \emph{not} a proof of equivariance of the graph convolution for a specific sampling. We simply observe that the graph Laplacian converges pointwisely toward an operator that commutes with rotation. To have the desired equality in eq..., we would need to prove a stronger result, for example uniform convergence (in the space of the operators).
\nati{Should we put this here?: }This does not hold in general \cite{belkin-spectral}. However, we believe that this is achievable for a bandlimited signals.
More importantly, the proof of ... in Appendix ... inspires us in the construction of our graph Laplacian. a) use gaussian kernel b) we need to adjust t with respect of n such that ...

@@

\subsection{Theoretical evidence of equivariance}

Given a sampling $x_0, \dots, x_{n-1}$ define $\sigma_i$ to be the patch of the surface of the sphere corresponding to the $i$-th point of the sampling, define $A_i$ to be its corresponding area and $d_i$ to be the radius of the smallest ball in $\mathbb R^3$ containing the $i$-th patch. Define $d^{(n)} := \max_{i=0, \dots, n}d_i$ and $A^{(n)}=\max_{i=0, \dots, n}A_i$.

\subsection{Equivariance error}

@@

\begin{definition}{}(Heat Kernel Graph Laplacian operator)\\
	\label{def:Heat Kernel Graph Laplacian operator}
	\text{Given a sampling $\{x_i\in\mathcal M\}_{i=0}^{n-1}$ of the manifold we define the \textbf{operator} }$L_n^t$ such that
	$$L_n^tf(y) := \frac{1}{n}\left[ \sum_{i=0}^{n-1} \exp{-\frac{||x_i-y||^2}{4t}}\left(f(y)-f(x_i)\right)\right]$$
\end{definition}
\begin{definition}{}(from \cite[Belkin et al. ]{Belkin:2005:TTF:2138147.2138189}Functional approximation to the Laplace-Beltrami operator)\\ \label{eq: my L^t} Let $\mu$ be the uniform probability measure on the manifold $\mathcal M$, where $\text{vol}(\mathcal M)$ is the volume of $\mathcal M$. We define the functional approximation to the Laplace-Beltrami operator to be the operator $L^t: L^{2}(\mathcal{M}) \rightarrow L^{2}(\mathcal{M})$ such that
	\label{def:Functional approximation to the Laplace-Beltrami operator}
	$$ L^tf(y) = \int_{\mathcal M}e^{-\frac{||y-x||^2}{4t}}\left(f(y)-f(x)\right)d\mu(x)$$
\end{definition}

@@

\begin{prop} 	Let $\mathcal M$ be a $k$-dimensional compact smooth manifold embedded in some euclidean space $\mathbb R^N$, and fix $p\in\mathcal M$. Let the data points $x_1, \dots, x_n$ be sampled form a uniform distribution on the manifold $\mathcal M$. Set $t_n=n^{-\frac{1}{k+2+\alpha}}$, for any $\alpha>0$ and let $f\in\mathcal C_\infty(\mathcal M)$. Then
	$$\frac{1}{t}\frac{1}{(4\pi t)^{k/2}} L^tf(p) \xrightarrow{t\to 0 } \frac{1}{\text{vol}(\mathcal M)}\triangle_{\mathcal M}f(p).$$
	\label{prop:3}
\end{prop}

@@

>>>>

% The plan for this section
% 0) Explain why the Laplacian is so important...
By definition, the graph convolution (on the sphere) is rotation equivariant to rotation, if and only if, it commutes with the rotation operator. 
As, in the context of this work, graph convolution can be performed by successive application of the graph Laplacian, i.e: $g(\b{L}) \b{f} = \left(\sum_{i=0}^K \alpha_i \b{L}^i\right) \b{f}$, if the graph Laplacian $\b{L}$ commutes with the rotation operator, then, by recursion, it will also commute with the convolution $g(\b{L})$.
Hence to simplify our analysis, we concentrate on the graph Laplacian.

% 0.5) Explain the sampling and metric
We first need to deal with the sampling problem. 
Take a sampling scheme $V=\{v_i\in\mathbb S^2, i=0, \dots, n\}$ of the sphere, a weighted undirected graph $G(V, E, \mathbf W)$, a signal $f: \mathbb S^2\to\mathbb R$ and its sampled representation $\mathbf f:\ f_i=f(v_i)$. Suppose that there exists a sampling operator $T_V: L^2(\mathbb S^2) \supset F\to \mathbb R^n,\  T_V(f) = \mathbf f$ defined on a suitable subspace $F$ of $L^2(\mathbb S^2)$ such that it is invertible, i.e., we can unambiguously reconstruct the function $f\in F$ from its sampled values $\mathbf f$. The existence of such subspace depends on the sampling scheme $V$, and its characterization is a common problem in signal processing \cite{Driscoll:1994:CFT:184069.184073}. Define the rotation operator $R(g), g\in SO(3)$: $R(g) f(\omega) = f\left(g^{-1} \omega\right)$.

We now want to understand how to set the edges and the weights of $G$ such that
\begin{equation} \label{eq:equivariance}
	T \R(g) T^{-1} \Omega_k T f = \Omega_k T R(g) f,
\end{equation}
i.e., the graph convolution $\Omega_k$ and any rotation $R(g)$ commute.

Verifying equation (\ref{eq:equivariance}) is really hard in practice, due to the fact that for almost all sampling schemes $V$ it is not known if there exists a space $F$ in which $T$ is invertible. A special case is the \textit{equiangular sampling} scheme where a sampling theorem holds \cite{Driscoll:1994:CFT:184069.184073}, and thus we have a closed form of $T^{-1}$. For sampling schemes where no such sampling theorem is available, there are implementations of discrete SHT to reconstruct a sampled signal $\mathbf f$, thus providing a way to approximate $T^{-1}$. Thanks to this we are able, for a given sampling, a given function $f$, a given rotation $g$, and a given kernel $k$, to compute the \textit{normalized equivariance error}
\begin{equation} \label{eq:equivariance error}
	E_{G}(f, g) = \left(\frac{ \norm {T R(g) T^{-1} \Omega_k Tf - \Omega_k T R(g) f}_{L^2(\mathbb R^2)}}{\norm {Tf}_{L^2(\mathbb R^2)}}\right)^2,
\end{equation}
where $T^{-1}$ is substituted with a discrete SHT in case $T$ is not invertible.
A measure of how equivariant a graph is with respect to rotations will then be given by the \textit{mean equivariance error}
\begin{equation} \label{eq:mean equivariance error}
	\overline E_G = \mathbb E_{f, g} \ E_G(f, g).
\end{equation}
In practice the expected value is obtained by averaging over a finite number of random functions and random rotations. The mean equivariance error $\overline E_G$ gives us an indication of how close the graph $G$ is from being equivariant to rotations. 
Of course result will depend on the sampling and on the graph construction.

% !) summary the work of Khasanova
\cite{khasanova2017graphomni} designed a discrete Laplacian that is explicitly intended to work on the sphere with the equiangular sampling. They consider the set $\mathcal G$ of all the possible graphs where each node is connected only to four of its nearest neighbours (North, Sud, West, East) and propose a weighting scheme $w_{ij}$ that minimizes the difference in the response to the polynomial spectral filter $\mathcal F = \mathbf L$ evaluated on images of the same object seen at different latitudes. In other words, they solve the minimization problem
\begin{equation} \label{eq:minimization frossard}
	\min_{W\in\mathcal G} \left|\mathcal{F}\left(\mathbf{y}\left(v_{ e}\right)\right)-\mathcal{F}\left(\mathbf{y}\left(v_{ i}\right)\right)\right|
\end{equation}
for the adjacency matrix $W$, where $\mathbf y(v_i)$ is the image of the object on the sphere centered on the vertex $v_i$, and $\mathcal F (\mathbf y(v_e))$ is the response of the filter at the vertex $v_e$ that lies at the same longitude of the vertex $v_i$ but on the equator. In their work they prove that the optimal weights solving the minimization problem (\ref{eq:minimization frossard}) are given by weights $w_{ij}$ inversely proportional to the Euclidean distance between vertices:
\begin{equation} \label{eq:frossard weights}
	w_{ij} = \frac{1}{\norm{x_i-x_j}}
\end{equation}

% 2) explain our approach
Our approach is different as we do not miminize eq \ref{eq:minimization frossard}. Instead we inspire ourselves with Belkin, which says that for \emph{uniform sampling} the graph Laplacian converges toward the Laplace-Beltrami. b) The Laplace Beltrami commute with rotation. The following theorem extend the work of Belkin and Nyiogi to fixed sampling \nati{(Including Healpix or not?)}. 
Given a sampling $V = \{x_0, \dots, x_{n-1}\}$ define $\sigma_i$ to be the patch of the surface of the sphere corresponding to $x_i$, define $A_i$ to be its corresponding area and $d_i$ to be the radius of the smallest ball in $\mathbb R^3$ containing $\sigma_i$. Define $d^{(n)} := \max_{i=0, \dots, n}d_i$ and $A^{(n)}=\max_{i=0, \dots, n}A_i$.

\begin{theorem}
	For a sampling $\mathcal P = \{x_i\in\mathbb S^2\}_{i=0}^{n-1}$ of the sphere that is equi area and such that $d^{(n)} \leq \frac{C}{\sqrt{n}}$, for all $f: \mathbb S^2 \rightarrow \mathbb R$ Lipschitz with respect to the euclidean distance in $\mathbb R^3$, for all $y\in\mathbb S^2$, there exists a sequence $t_n = n^\beta$ such that the rescaled Heat Kernel Graph Laplacian operator $\frac{|\mathbb S^2|}{4\pi t_n}L^t_n$ converges pointwise to the Laplace Beltrami operator on the sphere $\triangle_{\mathbb S^2}$  for $n\to\infty$:
	$$ \lim_{n\to\infty}\frac{|\mathbb S^2|}{4\pi t_n} L_n^{t_n}f(y) =  \triangle_{\mathbb S^2}f(y).$$
	\label{theo:pointwise convergence for a regular sampling}
\end{theorem}

This is \emph{not} a proof of equivariance of the graph convolution for a specific sampling. We simply observe that the graph Laplacian converges pointwisely toward an operator that commutes with rotation. To have the desired equality in eq..., we would need to prove a stronger result, for example uniform convergence (in the space of the operators).
\nati{Should we put this here?: }This does not hold in general \cite{belkin-spectral}. However, we believe that this is achievable for a bandlimited signals.
More importantly, the proof of theorem \ref{theo:pointwise convergence for a regular sampling} in Appendix ... inspires us in the construction of our graph Laplacian. a) use gaussian kernel b) we need to adjust t with respect of n such that ...

@@

@@

\begin{definition}{}(Heat Kernel Graph Laplacian operator)\\
	\label{def:Heat Kernel Graph Laplacian operator}
	\text{Given a sampling $\{x_i\in\mathcal M\}_{i=0}^{n-1}$ of the manifold we define the \textbf{operator} }$L_n^t$ such that
	$$L_n^tf(y) := \frac{1}{n}\left[ \sum_{i=0}^{n-1} \exp{-\frac{||x_i-y||^2}{4t}}\left(f(y)-f(x_i)\right)\right]$$
\end{definition}
\begin{definition}{} \cite{Belkin:2005:TTF:2138147.2138189}\\ \label{eq: my L^t} Let $\mu$ be the uniform probability measure on the manifold $\mathcal M$, and let $\text{vol}(\mathcal M)$ be the volume of $\mathcal M$. We define the functional approximation to the Laplace-Beltrami operator to be the operator $L^t: L^{2}(\mathcal{M}) \rightarrow L^{2}(\mathcal{M})$ such that
	\label{def:Functional approximation to the Laplace-Beltrami operator}
	$$ L^tf(y) = \int_{\mathcal M}e^{-\frac{||y-x||^2}{4t}}\left(f(y)-f(x)\right)d\mu(x)$$
\end{definition}

@@

\begin{prop} \cite{Belkin:2005:TTF:2138147.2138189} \\
Let $\mathcal M$ be a $k$-dimensional compact smooth manifold embedded in some euclidean space $\mathbb R^N$, and fix $p\in\mathcal M$. Let the data points $x_1, \dots, x_n$ be sampled form a uniform distribution on the manifold $\mathcal M$. Set $t_n=n^{-\frac{1}{k+2+\alpha}}$, for any $\alpha>0$ and let $f\in\mathcal C_\infty(\mathcal M)$. Then
	$$\frac{1}{t}\frac{1}{(4\pi t)^{k/2}} L^tf(p) \xrightarrow{t\to 0 } \frac{1}{\text{vol}(\mathcal M)}\triangle_{\mathcal M}f(p).$$
	\label{prop:3}
\end{prop}

@@

\subsection{MATERIAL FROM MARTINO'S MASTER PROJECT}

Given a sampling $x_0, \dots, x_{n-1}$ define $\sigma_i$ to be the patch of the surface of the sphere corresponding to the $i$th point of the sampling, define $A_i$ to be its corresponding area and $d_i$ to be the radius of the smallest ball in $\mathbb R^3$ containing the $i$-th patch. Define $d^{(n)} := \max_{i=0, \dots, n}d_i$ and $A^{(n)}=\max_{i=0, \dots, n}A_i$.

\subsection{MATERIAL FROM MARTINO'S MASTER PROJECT}
